---
output: github_document
  md_document:
    variant: markdown_github
---
# Exam Data Science
Student number 28776534

I created this project in an empty folder called 28776534 using the fmxdat::make_project(Open=T) command. 
General note: Most interpretations, especially detailed interpretations can be found in the pdf documents created with Texevier.

More detailed code explanations can be found in the codes for the functions. 

```{r}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
library(pacman)
p_load(tidyverse, lubridate, knitr, dplyr, readxl)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

Create folders + Texevier files for the 5 questions:

```{r}
Texevier::create_template(directory = "D:/Uni_Stellenbosch/Data_Science/28776534",
                          template_name = "Question1",
                          build_project = F, open_project = F)
Texevier::create_template(directory = "D:/Uni_Stellenbosch/Data_Science/28776534",
                          template_name = "Question2",
                          build_project = F, open_project = F)
Texevier::create_template(directory = "D:/Uni_Stellenbosch/Data_Science/28776534",
                          template_name = "Question3",
                          build_project = F, open_project = F)
Texevier::create_template(directory = "D:/Uni_Stellenbosch/Data_Science/28776534",
                          template_name = "Question4",
                          build_project = F, open_project = F)
Texevier::create_template(directory = "D:/Uni_Stellenbosch/Data_Science/28776534",
                          template_name = "Question5",
                          build_project = F, open_project = F)

```


# Question 1:

Initial Settings + Load Data 
```{r}
list.files('Question1/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

# As the Population_data was not provided in the data folder, I downloaded it from the website that was
#  mentioned in the txt.file. (https://data.world/bob-burggraaf/population-by-city-in-usa)
# Also as the files provided were RDS files and not csv, I changed that in the code.

Baby_Names <- readRDS("Question1/data/US_Baby_names/Baby_Names_By_US_State.rds")
charts <- readRDS("Question1/data/US_Baby_names/charts.rds")
#Population_data <- read_excel("Question1/data/US_Baby_names/Total_Population_By_City_Acs_2015_5_E_AgeSex.xlsx")
HBO_Titles <- readRDS("Question1/data/US_Baby_names/HBO_titles.rds")
HBO_Credits <- readRDS("Question1/data/US_Baby_names/HBO_credits.rds")


#Additional data source: Religion might also determine the baby name: 
# I found a data set listing typical regligious names in the data set "Hitchcocks Bible Dictionary"
library("httr")
GET("https://query.data.world/s/c7jlg53cop4hktqsqugcfeo53r2buh?dws=00000", write_disk(tf <- tempfile(fileext = ".xlsx")))
Bible_Names <- read_excel(tf) 

```


General insights into baby naming trends between 1910-2014 
- Top 25 Baby Names
- Spearman rank correlation - Baby names persistence over Time 
```{r}
# First extracting the top 25 baby names
source("Question1/code/Top_Baby_Names.R")
Top25BabyNames <- extract_top_n_names(data=Baby_Names, top_n=25)
Top25BabyNames

source("Question1/code/Spearman_Corr.R")
spearman_corr <- calculate_and_plot_correlations(data=Baby_Names) 
spearman_corr
```
Note:
Especially in the 1960s, female names do not seem to be correlated. However, until 1950, there
seems to be a correlation for both male and female baby names. Please find more interpretations in the PDF.


Factors influencing baby names
```{r}
#Looking for spikes in baby names 
source("Question1/code/Surges.R")
surges <- identify_year_on_year_surges(data=Baby_Names, percent_change=2000) 
surges
# interestingly, in 2012, there was an increase of 2010 % of the name Khaleesi (from 5 to 110)
# My theory is that this must be related to Game of Thrones (which was according to the HBO_Titles dataset released in 2012)
game_of_thrones_popularity <- HBO_Titles %>%
  filter(title == "Game of Thrones") %>%
  select(title, tmdb_popularity, release_year)
game_of_thrones_popularity

#Vizualizing the surges:
source("Question1/code/Visualize_surges.R")
visualization <- visualize_surges(data=surges, names_to_highlight=c("Neymar", "Krystle", "Khaleesi", "Mallory")) 
visualization
```
More detailed interpretations can be found in the PDF document for Question 1




# Question 2 Music Taste:

Initial Settings + Data Loading + First Cleaning
```{r}
list.files('Question2/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

#Data Loading
# Note that spotify dataset and billboard_100 were provided as RDS files.
coldplay <- read_csv("Question2/data/Coldplay_vs_Metallica/Coldplay.csv")
metallica <- read_csv("Question2/data/Coldplay_vs_Metallica/metallica.csv")
spotify <- readRDS("Question2/data/Coldplay_vs_Metallica/Broader_Spotify_Info.rds")
billboard_100 <- readRDS("Question2/data/Coldplay_vs_Metallica/charts.rds")

```
```{r}
# Filter out live recordings for direct comparisons
suppressWarnings({
coldplay_studio <- coldplay[!grepl('Live', coldplay$name, ignore.case = TRUE), ] %>%
    mutate(release_year = as.numeric(format(as.Date(release_date), "%Y")))

metallica_studio <- metallica[!grepl('Live', metallica$name, ignore.case = TRUE), ] %>%
        mutate(release_year = as.numeric(format(as.Date(release_date), "%Y")))

billboard_100_studio <- billboard_100[!grepl('Live', billboard_100$song, ignore.case = TRUE), ] 

spotify_studio <- spotify[!grepl('Live', spotify$name, ignore.case = TRUE), ] 
}) 

```


Comparing average popularity of the two bands over time
```{r}
library(ggplot2)
source("Question2/code/Popularity.R")
plot_popularity_over_time(data_coldplay=coldplay_studio, data_metallica=metallica_studio)
# Coldplay only exists since 1997 -> explains why graph for Coldplay starts so late
```

Popularity by Album for each band, Metallica and Coldplay
```{r}
source("Question2/code/Album_popularity_Metallica.R")
plot_Metallica <- create_boxplot_Metallica(data=metallica_studio, title="Popularity by album - Metallica")
plot_Metallica

source("Question2/code/Album_popularity_Coldplay.R")
plot_Coldplay <- create_boxplot_Coldplay(data=coldplay_studio, title="Popularity by album - Coldplay")
plot_Coldplay
```

How often were Coldplay and Metallica in the top 100 Billboard charts (excluding features) + comparison with band/singer that has most top 10 rankings?
```{r}
source("Question2/code/Billboard_Top_10.R")
plot_top10_mentions(data=billboard_100_studio, artist1 = "Coldplay", artist2 = "Metallica")

# As compared total appearances in top 100 for Coldplay and Metallica:
Coldplay_mentions <- billboard_100_studio %>%
    filter(artist == "Coldplay")
    
Metallica_mentions <- billboard_100_studio %>%
    filter(artist == "Metallica") 
    
Coldplay_total_count <- nrow(Coldplay_mentions)
Metallica_total_count <- nrow(Metallica_mentions)
Coldplay_total_count # 281 times ranked in top 100
Metallica_total_count # 197 times ranked in top 100
```
Finally, using spotify data set to compare and analyse how valence for the two bands developed over time
```{r}
source("Question2/code/Valence.R")
plot_valence <- plot_valence_over_time(data=spotify_studio)
plot_valence
```
Detailed interpretations for this can again be found in the PDF document



# Question 3 - Ukraine-Russia Conflict:

Initial settings
```{r}
list.files('Question3/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
```

Load Data sets:
```{r}
alloc <- read_csv("Question3/data/Ukraine_Aid/Financial Allocations.csv")
commit <- read_csv("Question3/data/Ukraine_Aid/Financial Commitments.csv")

# Merge the data sets based on country name and EU_membership status
aid_data_without_EU <-left_join(alloc, commit, by = c("Country", "EU member"), suffix = c("_allocation", "_commitment"))
aid_data <- full_join(alloc, commit, by = c("Country", "EU member"), suffix = c("_allocation", "_commitment"))
# note that EU (Commission and Council) only exists in the commit data set, which is why I used full_join to not 
#   drop that 
```

Having a general look at the relationship between GDP in 2021 and total aid allocations by country
```{r}
source("Question3/code/Comparison_Allocation_GDP.R")
# Function to compare bilateral allocations and commitments for EU members and analyze with GDP
Graph <- scatter_plot_fct(data=aid_data, title="Relationship between GDP in 2021 and Total Bilateral Allocations")
Graph
```


Comparing for EU member countries the commitments vs allocation (calculating the difference)
"Commitment gap"
```{r}
source("Question3/code/Comparison_Commitment_Allocation.R")
Comparison <- compare_eu_countries(data=aid_data, top_n=3)
Comparison
```
Should be 0 if countries allocated the same amount of money they actually commited to. For those countries
with negative numbers -> Commitments > Allocations 


Now looking at the difference, we might want to check again how the difference is correlated with GDP
```{r}
source("Question3/code/Comparison_Commitment_Allocation_GDP.R")
Comparison2 <- compare_diff_GDP(data=aid_data, title= "Difference between Allocations and Commitments vs GDP")
Comparison2
```
The higher GDP the more negative is the difference which means that countries with a higher GDP seem to have
made higher commitments than actual allocations. However, one must note, that there is a large variance with increasing GDP. Richest countries have the biggest commitment gap. 

Detailed interpretations can be found in the according pdf created with Texevier.



# Question 4 - Olympics:
Initial settings
```{r}
list.files('Question4/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

```

Load Data 
Note: The files that were given are RDS files.
```{r}
Winter <- readRDS("Question4/data/olympics/winter.rds")
Summer <- readRDS("Question4/data/olympics/summer.rds")
GDP <- readRDS("Question4/data/olympics/GDP.rds")
```


How has India faired in past summer Olympics compared to similarly sized economies, to
other emerging market economies and also select South American countries?
```{r}
#Merging the data (Summer and GDP)
Summer_GDP <- left_join(Summer, GDP, by = c("Country" = "Code"))
#create a vector with in total 10 emerging countries having a relatively similar GDP per Capita as India
#  from different continents/ can also be considered as emerging countries
emerging_country_vector <- c("IND", "KEN", "PAK", "UKR", "COL", "INA", "JAM", "ETH", "GEO", "RSA")
subset_Summer_GDP <- Summer_GDP %>%
  filter(Country %in% emerging_country_vector)
```
Since there are no observations for many Latin American countries, that have a relatively similar GDP as India,
I chose the following tow: Jamaica and Colombia. 
```{r}
source("Question4/code/India.R")
Plot <- Medals_India(data=subset_Summer_GDP)
Plot
```
India has thick green line - mainly won Hockey, which is a team sport -> which is why so low. 


Which countries have been most dominant in both Winter and Summer Olympics over
time? Maybe show some time-series analyses of a few countries side-by-side.
```{r}
source("Question4/code/Succesful_countries_Summer.R")
Plot_Summer <- plot_medals_summer(data=Summer)
Plot_Summer

source("Question4/code/Succesful_countries_Winter.R")
Plot_Winter <- plot_medals_winter(data=Winter)
Plot_Winter
```
There is probably a bias cause could not account for all team sports in the data. 


Which countries best punch above their weight when it comes to winning medals (however
you define this)?
Using the Population variable from the GDP data set

Winter Olympics
```{r}
Winter_GDP <- left_join(Winter, GDP, by = c("Country" = "Code"))
source("Question4/code/Winter_Gold.R")
gold_medals_per_capita_winter <- gold_medals_winter(data=Winter_GDP)
gold_medals_per_capita_winter
```

Summer Olympics
```{r}
source("Question4/code/Summer_Gold.R")
gold_medals_per_capita_summer <- gold_medals_summer(data=Summer_GDP)
gold_medals_per_capita_summer
```


Which is your personal favourite event at the Olympics? Show some interesting analyses
on past winners / countries related to your chosen event
I generally prefer Winter Olympics -Alpine Skiing (Slalom) 

```{r}
source("Question4/code/Alpine_Skiing.R")
Top_Athletes <- Alpine_Skiing(data=Winter)
Top_Athletes
```

Detailed descriptions and interpretations can be found in the pdf



# Question 5 - SQL database: 
Initial settings + Renviron
```{r}
list.files('Question5/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
usethis::edit_r_environ()
# restart R

```
# Loading data + libraries + settings
```{r}
library(tidyverse)
library(dbbasic)
library(tidytext)
library(quanteda)
library(topicmodels)
list.files('Question5/code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
conn <- db_connect(db = "psql_datascience")
DBI::dbDisconnect(conn)

retailer <- db_query("SELECT * FROM retailer;", db = "psql_datascience")  
statssa <- db_query("SELECT * FROM statssa;", db = "psql_datascience")  

```

Data Cleaning + Filtering 
```{r}
# vector with braai ingredients
ingredients <- c("bread", "cheese", "margarine", "tomatoes", "onions", "salt", "chutney")

retailer_filtered <- retailer %>% 
  filter(category %in% ingredients) 

statssa_filtered <- statssa %>% 
  filter(grepl("white bread|cheddar|margarine|tomatoes|onions|salt|chutney", type, ignore.case = TRUE)) %>%
    mutate(ingredient = str_to_lower(str_extract(type, "white bread|cheddar|margarine|tomatoes|onions|salt|chutney|Tomatoes|Onions|Salt|Chutney|White bread|Cheddar|Margarine"))) %>%
    filter(date >= as.Date("2020-03-29"))  
# Filter for observations starting from 2020-03-29 to align with retailer data!!
## also made everything to lower case so it is like category variable 
```


Retailer Index: 
# first I calculated the index (using monthly price averages). The base price is the first months average, as the data would be too noisy if used on a daily basis. 
Then I plotted the index over time.
More details can be found in the code itself. 

```{r}
source("Question5/code/Retailer.R")
retailer_plot <- retailer_plot_fct(data=retailer_filtered)
retailer_plot

```

Stats SA Analysis
# first I calculated the index (using monthly price averages). The base price is the first months average, as the data would be too noisy if used on a daily basis. 
Then I plotted the index over time. 
More details can be found in the code itself. 
```{r}
source("Question5/code/StatsSA.R")
statssa_plot <- statssa_plot_fct(data=statssa_filtered)
statssa_plot
```


B - basic summary statistics and comparison of the two data sets:

Having a look at summary statistics, in particular mean prices of different products in the two different data sets
```{r}
source("Question5/code/summary_stats.R")
plot <- plot_summary_stats(data1=retailer_filtered, data2=statssa_filtered)
plot
```

Having a look at average monthly tomato prices in the two different data sets: 

```{r}
source("Question5/code/tomatoe_prices.R")
plot <- plot_monthly_tomato_prices(data1=retailer_filtered, data2=statssa_filtered)
plot
```
Note!
Retailer calculates tomato prices per kg 
no quantity mentioned for statssa data




# end of ReadMe file